Galactic Chemical Evolution Models Favor an Extended Type Ia Supernova Delay-Time Distribution

### 2024-04-03 src/tex/ms.tex
[showyourwork] overleaf sync

% Abstract from AAS 241
\begin{abstract}
    Type Ia supernovae (SNe Ia) produce most of the Fe-peak elements in the Universe and therefore are a crucial ingredient in galactic chemical evolution models. SNe Ia do not explode immediately after star formation, and the delay-time distribution (DTD) has not been definitively determined by supernova surveys or theoretical models. Because the DTD also affects the relationship among age, [Fe/H], and \aFe \todo{([O/Fe]?)} in chemical evolution models, comparison with observations of stars in the Milky Way is an important consistency check for any proposed DTD. We implement several popular forms of the DTD in combination with multiple star formation histories for the Milky Way in %one-zone and 
    multi-zone chemical evolution models which include radial stellar migration.
    % using the Versatile Integrator for Chemical Evolution (\vice). \vice includes a prescription for radial stellar migration in multi-zone models. 
    We compare our predicted interstellar medium abundance tracks, stellar abundance distributions, and stellar age distributions to the final data release of the Apache Point Observatory Galactic Evolution Experiment (APOGEE). We find that the DTD has the largest effect on the \aFe distribution: a DTD with more prompt SNe Ia produces a stellar abundance distribution that is skewed toward a lower \aFe ratio.
    % whereas a DTD with more delayed SNe Ia produces a distribution that is skewed toward higher \aFe.
    While the DTD alone cannot explain the observed bimodality in the \aFe distribution, in combination with an appropriate star formation history it affects the goodness of fit between the simulated and observed high-$\alpha$ sequence. Our model results favor an extended DTD with fewer prompt SNe Ia than the fiducial $t^{-1}$ power law.
\end{abstract}

@@

% [Theoretical DTDs]
% Uncertainties in the progenitor channels of SNe Ia translate into uncertainties in the shape of the DTD. 
% By formulating analytical DTDs for several progenitor models, \citet{Greggio2005-AnalyticalRates} found that the SD channel tends to produce a steeper DTD function than the DD channel, and the median delay time (by which 50\% of SNe Ia have exploded) can vary by a factor of 10 depending on the chosen progenitor scenario and its particular parameters.
As a result of the uncertainty regarding SN Ia progenitors, theoretical models have yet to converge on a single prediction for the DTD.
For the DD channel, assumptions about the distribution of post-common envelope separations and the rate of gravitational wave inspiral suggest a broad $\sim t^{-1}$ DTD at long delay times ($\gtrsim 1$ Gyr), but at short delays ($\lesssim 1$ Gyr) the rate is limited by the need to produce two WDs \citep[see][]{Maoz2014-Review}.
Triple or higher-order progenitor systems could also produce a $t^{-1}$ DTD \citep{Fang2018-QuadrupleSystems,Rajamuthukumar2023-TripleEvolution}.
The DTD which would result from the SD channel depends greatly on the assumptions of binary population synthesis, but in general is expected to cover a narrower range of delay times and may feature a steep exponential cutoff at the long end \citep[e.g.,][]{Greggio2005-AnalyticalRates}.
% More recently, \citet{Rajamuthukumar2023-TripleEvolution} computed a DTD from the evolution of triple system progenitors. [Expand, connect binary evolution to delay times, look for other theory papers]

@@

This paper presents the first\todo{(?)} comprehensive look at the DTD in a multi-zone GCE model that can qualitatively reproduce the observed abundance structure of the Milky Way disk. A multi-zone approach allows for a radially-dependent parameterization of the SFH, outflows, stellar migration, and abundance gradient which can better match observations across the Galactic disk. We evaluate a selection of DTDs from the literature with multiple SFHs and a prescription for radial stellar migration in the Versatile Integrator for Chemical Evolution \citep[\vice;][]{JohnsonWeinberg2020-Starbursts}. In Section \ref{sec:methods}, we present our models for the DTD and SFH and describe our observational sample. In Section \ref{sec:onezone-results}, we detail our one-zone chemical evolution models and present results. In Section \ref{sec:multizone-results}, we present the results of our multi-zone models and compare to observations. In Section \ref{sec:discussion}, we discuss the implications for the DTD and future surveys. In Section \ref{sec:conclusions}, we summarize our conclusions.

@@

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models. We omit some parameters that are unchanged from \citetalias{Johnson2021-Migration}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone model parameters
        R_{\rm gal}     & [0, 20] kpc   & \ref{sec:multizone-results} & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & \ref{sec:multizone-results} & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\tau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration distance \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\tau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & \ref{sec:multizone-results} & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & \ref{sec:multizone-results} & Maximum simulation time \\
        n               & 8         & \ref{sec:multizone-results} & Number of stellar populations formed per ring per time-step \\
        R_{\rm SF}      & 15.5 kpc  & \ref{sec:multizone-results} & Maximum radius of star formation \\
        M_{g,0}   & 0         & \ref{sec:sfh}     & Initial gas mass \\
        \dot M_r    & continuous    & \ref{sec:multizone-results} & Recycling rate \citep[][Equation 2]{JohnsonWeinberg2020-Starbursts} \\
        \hline
        % DTD
        R_{\rm Ia}(t)   & Equation \ref{eq:dtd-function}    & \ref{sec:dtd-models}  & delay-time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        N_{\rm Ia}/M_\star  & $2.2\times10^{-3}$ M$_\odot^{-1}$ & \ref{sec:yields}  & SNe Ia per unit mass of stars formed \citep{MaozMannucci2012-SNeIaReview} \\
        \hline
        % Nucleosynthetic yields
        y_{\rm O}^{\rm CC}  & 0.015     & \ref{sec:yields}  & CCSN yield of O    \\
        y_{\rm Fe}^{\rm CC} & 0.0012    & \ref{sec:yields}  & CCSN yield of Fe   \\
        y_{\rm O}^{\rm Ia}  & 0         & \ref{sec:yields}  & SN Ia yield of O       \\
        y_{\rm Fe}^{\rm Ia} & 0.00214   & \ref{sec:yields}  & SN Ia yield of Fe \\
        \hline
        % Star formation histories
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & \ref{sec:sfh} & Time-dependence of the inside-out SFR \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & \ref{sec:sfh} & Time-dependence of the late-burst SFR \\
        \tau_{\rm rise}             & 2 Gyr     & \ref{sec:sfh} & SFR rise timescale for inside-out and early-burst models \\
        \tau_{\rm EB}(t)          & Equation \ref{eq:earlyburst-taustar}  & \ref{sec:sfh}   & Time-dependence of the early-burst SFE timescale \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & \ref{sec:sfh} & Time-dependence of the early-burst infall rate \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & \ref{sec:sfh} & Time-dependence of the two-infall infall rate \\
        \hline
        % One-zone parameters
        \tau_\star                    & 2 Gyr & \ref{sec:onezone-results} & SFE timescale in one-zone models \\
        \eta(R_{\rm gal}=8\,{\rm kpc})  & 2.15  & \ref{sec:onezone-results} & Outflow mass-loading factor at the solar annulus \\
        \tau_{\rm sfh}(R_{\rm gal}=8\,{\rm kpc})    & 15.1 Gyr  & \ref{sec:sfh} & SFH timescale at the solar annulus \\
    \enddata
\end{deluxetable*}
\vspace{-24pt}

@@

Following the formalism of \citet{Weinberg2017-ChemicalEquilibrium}, the rate of Fe contribution to the ISM from SNe Ia is 
\begin{equation}
    \dot M_{\rm Fe}^{\rm Ia} = y_{\rm Fe}^{\rm Ia} \langle \dot M_\star\rangle_{\rm Ia} \\
\end{equation}
where $\langle \dot M_\star\rangle_{\rm Ia}$ is the time-averaged SFR weighted by the DTD and $y_{\rm Fe}^{\rm Ia}$ is the Fe yield of SNe Ia. \citet{Weinberg2017-ChemicalEquilibrium} show in their Appendix A that
\begin{equation}
    \langle \dot M_\star\rangle_{\rm Ia} \equiv \frac{\int_0^t \dot M_\star(t')R_{\rm Ia}(t-t')dt'}{\int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt'}
    \label{eq:weighted-sfr}
\end{equation}
where %$\dot M_\star(t)$ is the SFR at time $t$, 
%$R_{\rm Ia}(t) dt$ is the number of SNe Ia which explode in the time interval $(t, t+dt)$ per unit mass of stars formed at $t=0$, 
$R_{\rm Ia}$ is the DTD in units of ${\rm M}_\odot^{-1}\,{\rm yr}^{-1}$,
$t_D$ is the minimum SN Ia delay time, and $t_{\rm max}$ is the maximum simulation time. The denominator of Equation \ref{eq:weighted-sfr} is therefore equal to $N_{\rm Ia}/M_\star$, the total number of SNe Ia per M$_{\odot}$ of stars formed.

@@

We explore five different functional forms for the DTD: a two-population model, a single power-law, an exponential, a broken power-law with an initially flat plateau, and a model computed from triple-system dynamics. We also investigate one or two useful variations of the input parameters for each functional form. Figure \ref{fig:dtds} presents a selection of these DTDs, and Table \ref{tab:dtds} summarizes all of our DTDs and their parameters. We use simple forms rather than simulated physical or analytical models of SNe Ia for the sake of decreased computational time and easier interpretation of the model predictions. Physically-motivated models of the DTD must contend with many unknown or poorly-constrained parameters, so our simplified forms have the advantage of reducing the number of free parameters. In Appendix \ref{app:analytical-dtds}, we show that a few of our simple forms adequately approximate the more complete analytic models of \citet{Greggio2005-AnalyticalRates}.

@@

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$:
\begin{equation}
    f_{\rm Ia}^{\rm exp}(t) = e^{-t/\tau}.
    \label{eq:exponential-dtd}
\end{equation}
This model allows analytic solutions to the abundances as a function of time for some SFHs, making it a popular choice. \citet{Schonrich2009-RadialMixing} and \citet{Weinberg2017-ChemicalEquilibrium} both assumed an exponential DTD with a timescale $\tau=1.5$ Gyr. We show in Appendix \ref{app:analytical-dtds} that this is an adequate approximation for the analytical SD DTD from \citet{Greggio2005-AnalyticalRates}. 
However, there is less observational support for an exponential DTD. \citet{Strolger2020-ExponentialDTD}, fitting to the cosmic SFH and SFHs from field galaxies, found a range of exponential-like solutions with timescales $\sim 1.5 - 6$ Gyr.
In this paper, we investigate timescales $\tau=1.5$ and 3 Gyr.

\paragraph{Plateau} A modification of the power-law in which the DTD ``plateaus'' for a duration $W$ before declining:
\begin{equation}
    f_{\rm Ia}^{\rm plat}(t) =
    \begin{cases}
        1, & t < W \\
        (t/W)^\alpha, & t \ge W.
    \end{cases}
    \label{eq:plateau-dtd}
\end{equation}
% where $B_W=[(W-t_{\rm min}) + \frac{t_{\rm max}^{\alpha+1} - W^{\alpha+1}}{(\alpha+1)W^\alpha}]^{-1}$ is the normalization coefficient.
Our primary motivation is to consider a model which matches observations at delay times beyond a few Gyr, where the DTD is best constrained, but with a smaller fraction of prompt ($\lesssim 100$ Myr) SNe Ia than the single power law.
To our knowledge, this form of the DTD has not been considered in previous GCE models. We show in Appendix \ref{app:analytical-dtds} that this form can approximate the more complicated analytical DD DTDs from \citet{Greggio2005-AnalyticalRates}. We investigate the cases $W=0.3$ Gyr and $W=1$ Gyr, taking $\alpha=-1.1$ for all plateau models.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/star_formation_histories.pdf}
    \caption{The surface densities of star formation $\dot \Sigma_\star$ (first row), gas infall $\dot \Sigma_{\rm in}$ (second row), and gas mass $\Sigma_g$ (third row), and the SFE timescale $\tau_\star$ (fourth row) as functions of simulation time for our four model SFHs: inside-out (first column; see Equation \ref{eq:insideout-sfh}), late-burst (second column; see Equation \ref{eq:lateburst-sfh}), early-burst (third column; see Equations \ref{eq:earlyburst-ifr} and \ref{eq:earlyburst-taustar}), and two-infall (fourth column; see Equation \ref{eq:twoinfall-ifr}). In each panel, we plot curves for the model zones which have inner radii at 4 kpc (yellow), 6 kpc (orange), 8 kpc (red), 10 kpc (violet), 12 kpc (indigo), and 14 kpc (blue).}
    \label{fig:sfhs}
    \script{star_formation_histories.py}
\end{figure*}

@@

\paragraph{Two-infall} First proposed by \citet{Chiappini1997-TwoInfall}, this model parameterizes the infall rate as two successive, exponentially declining bursts to explain the origin of the high- and low-$\alpha$ disk populations:
\begin{equation}
    \label{eq:twoinfall-ifr}
    f_{\rm TI}(t|R_{\rm gal}) = N_1(R_{\rm gal}) e^{-t/\tau_1} + N_2(R_{\rm gal}) e^{-(t-t_{\rm on})/\tau_2},
\end{equation}
% In this model, the first infall produces the thick disk and the second infall produces the thin disk. 
where $N_1$ and $N_2$ are the normalizations of the first and second infall, respectively, and $t_{\rm on}$ is the onset time of the second infall.
The normalization ratio $N_2/N_1$ is calculated so that the thick-to-thin-disk surface density ratio $f_\Sigma(R)=\Sigma_2(R)/\Sigma_1(R)$ is given by
\begin{equation}
    f_\Sigma(R) = f_\Sigma(0) e^{R(1/R_2 - 1/R_1)}.
\end{equation}
Following \citet{BlandHawthornGerhard2016-MilkyWayReview}, we adopt values for the thick disk scale radius $R_1=2.0$ kpc, thin disk scale radius $R_2=2.5$ kpc, and $f_\Sigma(0)=0.27$.
We note that most previous studies which use the two-infall model \citep[e.g.,][]{Chiappini1997-TwoInfall,Matteucci2006-BimodalDTDConsequences,Matteucci2009-DTDModels,Spitoni2021-TwoInfall} do not consider gas outflows and instead adjust the nucleosynthetic yields to reproduce the solar abundance. We adopt radially-dependent outflows as in \citetalias{Johnson2021-Migration} (see their Section 2.4 for details) for all our SFHs, including two-infall. We discuss the implications of this difference in Section \ref{sec:two-infall-discussion}.

@@

Similar trends can be seen when adjusting the timescale of the exponential DTD, as shown in the middle panel of Figure \ref{fig:onezone-threepanel}. Here, the knee is not a sharp feature associated with the onset of SNe Ia as in the power-law case, but rather a gentle curve in the abundance track around a simulation time of 1 Gyr. Doubling of the timescale from 1.5 Gyr to 3 Gyr raises the [O/Fe] abundance ratio at $t=1$ Gyr by $\sim0.05$ dex and at $t=3$ Gyr by $\sim0.1$ dex, but the equilibrium abundance reached at $t=10$ Gyr remains unchanged. A longer exponential timescale also produces a broader [O/Fe] DF with more high-$\alpha$ stars, but the distribution is still unimodal.
The effect on the MDF is slightly more pronounced than the power-law case, with longer timescales skewing to lower [Fe/H] values.

@@

We run a total of multi-zone models with all combinations of our eight DTDs and four SFHs, for a total of 32. In the following subsections, we present the stellar abundance and age distributions from the multi-zone models and compare to APOGEE data from across the Galactic disk.

@@

The abundance tracks of all four model SFHs initially follow a similar path of decreasing [O/Fe] with increasing [Fe/H]. The ISM abundance ratios of the inside-out model change monotonically over the entire simulation time. The stellar abundance distribution at both low- and high-[O/Fe] is composed of stars with a wide range of birth $R_{\rm gal}$. In particular, we observe a high-[O/Fe] ``ridge'' on the high-metallicity side of the ISM track, which is composed of stellar populations from the inner Galaxy.

@@

\begin{figure}
    \centering
    \includegraphics{figures/ofe_feh_twoinfall.pdf}
    \caption{The [O/Fe]--[Fe/H] plane for multiple Galactocentric regions from the model with the inside-out SFH and plateau ($W=0.3$ Gyr) DTD. The two columns of panels contain stars in different bins of $R_{\rm gal}$, and each row contains stars from a different bin of $|z|$. The contents of each panel is as described in Figure \ref{fig:ofe-feh-sfh}.}
    \label{fig:ofe-feh-twoinfall}
    \script{ofe_feh_twoinfall.py}
\end{figure}

>>>>

% Abstract from AAS 241
\begin{abstract}
    Type Ia supernovae (SNe Ia) produce most of the Fe-peak elements in the Universe and therefore are a crucial ingredient in galactic chemical evolution models. SNe Ia do not explode immediately after star formation, and the delay-time distribution (DTD) has not been definitively determined by supernova surveys or theoretical models. Because the DTD also affects the relationship among age, [Fe/H], and \aFe in chemical evolution models, comparison with observations of stars in the Milky Way is an important consistency check for any proposed DTD. We implement several popular forms of the DTD in combination with multiple star formation histories for the Milky Way in %one-zone and 
    multi-zone chemical evolution models which include radial stellar migration.
    % using the Versatile Integrator for Chemical Evolution (\vice). \vice includes a prescription for radial stellar migration in multi-zone models. 
    We compare our predicted interstellar medium abundance tracks, stellar abundance distributions, and stellar age distributions to the final data release of the Apache Point Observatory Galactic Evolution Experiment (APOGEE). We find that the DTD has the largest effect on the \aFe distribution: a DTD with more prompt SNe Ia produces a stellar abundance distribution that is skewed toward a lower \aFe ratio.
    % whereas a DTD with more delayed SNe Ia produces a distribution that is skewed toward higher \aFe.
    While the DTD alone cannot explain the observed bimodality in the \aFe distribution, in combination with an appropriate star formation history it affects the goodness of fit between the simulated and observed high-$\alpha$ sequence. Our model results favor an extended DTD with fewer prompt SNe Ia than the fiducial $t^{-1}$ power law.
\end{abstract}

@@

% [Theoretical DTDs]
% Uncertainties in the progenitor channels of SNe Ia translate into uncertainties in the shape of the DTD. 
% By formulating analytic DTDs for several progenitor models, \citet{Greggio2005-AnalyticalRates} found that the SD channel tends to produce a steeper DTD function than the DD channel, and the median delay time (by which 50\% of SNe Ia have exploded) can vary by a factor of 10 depending on the chosen progenitor scenario and its particular parameters.
As a result of the uncertainty regarding SN Ia progenitors, theoretical models have yet to converge on a single prediction for the DTD.
For the DD channel, assumptions about the distribution of post-common envelope separations and the rate of gravitational wave inspiral suggest a broad $\sim t^{-1}$ DTD at long delay times ($\gtrsim 1$ Gyr), but at short delays ($\lesssim 1$ Gyr) the rate is limited by the need to produce two WDs \citep[see][]{Maoz2014-Review}.
Triple or higher-order progenitor systems could also produce a $t^{-1}$ DTD \citep{Fang2018-QuadrupleSystems,Rajamuthukumar2023-TripleEvolution}.
The DTD which would result from the SD channel depends greatly on the assumptions of binary population synthesis, but in general is expected to cover a narrower range of delay times and may feature a steep exponential cutoff at the long end \citep[e.g.,][]{Greggio2005-AnalyticalRates}.
% More recently, \citet{Rajamuthukumar2023-TripleEvolution} computed a DTD from the evolution of triple system progenitors. [Expand, connect binary evolution to delay times, look for other theory papers]

@@

This paper presents a comprehensive look at the DTD in a multi-zone GCE model that can qualitatively reproduce the observed abundance structure of the Milky Way disk. A multi-zone approach allows for a radially-dependent parameterization of the SFH, outflows, stellar migration, and abundance gradient which can better match observations across the Galactic disk. We evaluate a selection of DTDs from the literature with multiple SFHs and a prescription for radial stellar migration in the Versatile Integrator for Chemical Evolution \citep[\vice;][]{JohnsonWeinberg2020-Starbursts}. In Section \ref{sec:methods}, we present our models for the DTD and SFH and describe our observational sample. In Section \ref{sec:onezone-results}, we detail our one-zone chemical evolution models and present results. In Section \ref{sec:multizone-results}, we present the results of our multi-zone models and compare to observations. In Section \ref{sec:discussion}, we discuss the implications for the DTD and future surveys. In Section \ref{sec:conclusions}, we summarize our conclusions.

@@

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models. We omit some parameters that are unchanged from \citetalias{Johnson2021-Migration}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone model parameters
        R_{\rm gal}     & [0, 20] kpc   & \ref{sec:multizone-results} & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & \ref{sec:multizone-results} & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\tau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration distance \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\tau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & \ref{sec:multizone-results} & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & \ref{sec:multizone-results} & Disk lifetime \\
        n               & 8         & \ref{sec:multizone-results} & Number of stellar populations formed per ring per time-step \\
        R_{\rm SF}      & 15.5 kpc  & \ref{sec:multizone-results} & Maximum radius of star formation \\
        M_{g,0}   & 0         & \ref{sec:sfh}     & Initial gas mass \\
        \dot M_r    & continuous    & \ref{sec:multizone-results} & Recycling rate \citep[][Equation 2]{JohnsonWeinberg2020-Starbursts} \\
        \hline
        % DTD
        R_{\rm Ia}(t)   & Equation \ref{eq:dtd-function}    & \ref{sec:dtd-models}  & delay-time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        N_{\rm Ia}/M_\star  & $2.2\times10^{-3}$ M$_\odot^{-1}$ & \ref{sec:yields}  & SNe Ia per unit mass of stars formed \citep{MaozMannucci2012-SNeIaReview} \\
        \hline
        % Nucleosynthetic yields
        y_{\rm O}^{\rm CC}  & 0.015     & \ref{sec:yields}  & CCSN yield of O    \\
        y_{\rm Fe}^{\rm CC} & 0.0012    & \ref{sec:yields}  & CCSN yield of Fe   \\
        y_{\rm O}^{\rm Ia}  & 0         & \ref{sec:yields}  & SN Ia yield of O       \\
        y_{\rm Fe}^{\rm Ia} & 0.00214   & \ref{sec:yields}  & SN Ia yield of Fe \\
        \hline
        % Star formation histories
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & \ref{sec:sfh} & Time-dependence of the inside-out SFR \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & \ref{sec:sfh} & Time-dependence of the late-burst SFR \\
        \tau_{\rm rise}             & 2 Gyr     & \ref{sec:sfh} & SFR rise timescale for inside-out and early-burst models \\
        \tau_{\rm EB}(t)          & Equation \ref{eq:earlyburst-taustar}  & \ref{sec:sfh}   & Time-dependence of the early-burst SFE timescale \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & \ref{sec:sfh} & Time-dependence of the early-burst infall rate \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & \ref{sec:sfh} & Time-dependence of the two-infall infall rate \\
        \hline
        % One-zone parameters
        \tau_\star                    & 2 Gyr & \ref{sec:onezone-results} & SFE timescale in one-zone models \\
        \eta(R_{\rm gal}=8\,{\rm kpc})  & 2.15  & \ref{sec:onezone-results} & Outflow mass-loading factor at the solar annulus \\
        \tau_{\rm sfh}(R_{\rm gal}=8\,{\rm kpc})    & 15.1 Gyr  & \ref{sec:sfh} & SFH timescale at the solar annulus \\
    \enddata
\end{deluxetable*}
\vspace{-24pt}

@@

Following the formalism of \citet{Weinberg2017-ChemicalEquilibrium}, the rate of Fe contribution to the ISM from SNe Ia is 
\begin{equation}
    \dot M_{\rm Fe}^{\rm Ia} = y_{\rm Fe}^{\rm Ia} \langle \dot M_\star\rangle_{\rm Ia} \\
\end{equation}
where $\langle \dot M_\star\rangle_{\rm Ia}$ is the time-averaged SFR weighted by the DTD and $y_{\rm Fe}^{\rm Ia}$ is the Fe yield of SNe Ia. \citet{Weinberg2017-ChemicalEquilibrium} show in their Appendix A that
\begin{equation}
    \langle \dot M_\star\rangle_{\rm Ia} \equiv \frac{\int_0^t \dot M_\star(t')R_{\rm Ia}(t-t')dt'}{\int_{t_D}^{t_{\rm max}} R_{\rm Ia}(t')dt'}
    \label{eq:weighted-sfr}
\end{equation}
where %$\dot M_\star(t)$ is the SFR at time $t$, 
%$R_{\rm Ia}(t) dt$ is the number of SNe Ia which explode in the time interval $(t, t+dt)$ per unit mass of stars formed at $t=0$, 
$R_{\rm Ia}$ is the DTD in units of ${\rm M}_\odot^{-1}\,{\rm yr}^{-1}$,
$t_D$ is the minimum SN Ia delay time, and $t_{\rm max}$ is the lifetime of the disk. The denominator of Equation \ref{eq:weighted-sfr} is therefore equal to $N_{\rm Ia}/M_\star$, the total number of SNe Ia per M$_{\odot}$ of stars formed.

@@

We explore five different functional forms for the DTD: a two-population model, a single power-law, an exponential, a broken power-law with an initially flat plateau, and a model computed from triple-system dynamics. We also investigate one or two useful variations of the input parameters for each functional form. Figure \ref{fig:dtds} presents a selection of these DTDs, and Table \ref{tab:dtds} summarizes all of our DTDs and their parameters. We use simple forms rather than simulated physical or analytic models of SNe Ia for the sake of decreased computational time and easier interpretation of the model predictions. Physically-motivated models of the DTD must contend with many unknown or poorly-constrained parameters, so our simplified forms have the advantage of reducing the number of free parameters. In Appendix \ref{app:analytic-dtds}, we show that a few of our simple forms adequately approximate the more complete analytic models of \citet{Greggio2005-AnalyticalRates}.

@@

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$:
\begin{equation}
    f_{\rm Ia}^{\rm exp}(t) = e^{-t/\tau}.
    \label{eq:exponential-dtd}
\end{equation}
This model allows analytic solutions to the abundances as a function of time for some SFHs, making it a popular choice. \citet{Schonrich2009-RadialMixing} and \citet{Weinberg2017-ChemicalEquilibrium} both assumed an exponential DTD with a timescale $\tau=1.5$ Gyr. We show in Appendix \ref{app:analytic-dtds} that this is an adequate approximation for the analytic SD DTD from \citet{Greggio2005-AnalyticalRates}. 
However, there is less observational support for an exponential DTD. \citet{Strolger2020-ExponentialDTD}, fitting to the cosmic SFH and SFHs from field galaxies, found a range of exponential-like solutions with timescales $\sim 1.5 - 6$ Gyr.
In this paper, we investigate timescales $\tau=1.5$ and 3 Gyr.

\paragraph{Plateau} A modification of the power-law in which the DTD ``plateaus'' for a duration $W$ before declining:
\begin{equation}
    f_{\rm Ia}^{\rm plat}(t) =
    \begin{cases}
        1, & t < W \\
        (t/W)^\alpha, & t \ge W.
    \end{cases}
    \label{eq:plateau-dtd}
\end{equation}
% where $B_W=[(W-t_{\rm min}) + \frac{t_{\rm max}^{\alpha+1} - W^{\alpha+1}}{(\alpha+1)W^\alpha}]^{-1}$ is the normalization coefficient.
Our primary motivation is to consider a model which matches observations at delay times beyond a few Gyr, where the DTD is best constrained, but with a smaller fraction of prompt ($\lesssim 100$ Myr) SNe Ia than the single power law.
To our knowledge, this form of the DTD has not been considered in previous GCE models. We show in Appendix \ref{app:analytic-dtds} that this form can approximate the more complicated analytic DD DTDs from \citet{Greggio2005-AnalyticalRates}. We investigate the cases $W=0.3$ Gyr and $W=1$ Gyr, taking $\alpha=-1.1$ for all plateau models.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/star_formation_histories.pdf}
    \caption{The surface densities of star formation $\dot \Sigma_\star$ (first row), gas infall $\dot \Sigma_{\rm in}$ (second row), and gas mass $\Sigma_g$ (third row), and the SFE timescale $\tau_\star$ (fourth row) as functions of time for our four model SFHs: inside-out (first column; see Equation \ref{eq:insideout-sfh}), late-burst (second column; see Equation \ref{eq:lateburst-sfh}), early-burst (third column; see Equations \ref{eq:earlyburst-ifr} and \ref{eq:earlyburst-taustar}), and two-infall (fourth column; see Equation \ref{eq:twoinfall-ifr}). In each panel, we plot curves for the model zones which have inner radii at 4 kpc (yellow), 6 kpc (orange), 8 kpc (red), 10 kpc (violet), 12 kpc (indigo), and 14 kpc (blue).}
    \label{fig:sfhs}
    \script{star_formation_histories.py}
\end{figure*}

@@

\paragraph{Two-infall} First proposed by \citet{Chiappini1997-TwoInfall}, this model parameterizes the infall rate as two successive, exponentially declining bursts to explain the origin of the high- and low-$\alpha$ disk populations:
\begin{equation}
    \label{eq:twoinfall-ifr}
    f_{\rm TI}(t|R_{\rm gal}) = N_1(R_{\rm gal}) e^{-t/\tau_1} + N_2(R_{\rm gal}) e^{-(t-t_{\rm on})/\tau_2},
\end{equation}
% In this model, the first infall produces the thick disk and the second infall produces the thin disk. 
where $N_1$ and $N_2$ are the normalizations of the first and second infall, respectively, and $t_{\rm on}$\todo{=4 Gyr? (citations)} is the onset time of the second infall.
The normalization ratio $N_2/N_1$ is calculated so that the thick-to-thin-disk surface density ratio $f_\Sigma(R)=\Sigma_2(R)/\Sigma_1(R)$ is given by
\begin{equation}
    f_\Sigma(R) = f_\Sigma(0) e^{R(1/R_2 - 1/R_1)}.
\end{equation}
Following \citet{BlandHawthornGerhard2016-MilkyWayReview}, we adopt values for the thick disk scale radius $R_1=2.0$ kpc, thin disk scale radius $R_2=2.5$ kpc, and $f_\Sigma(0)=0.27$.
We note that most previous studies which use the two-infall model \citep[e.g.,][]{Chiappini1997-TwoInfall,Matteucci2006-BimodalDTDConsequences,Matteucci2009-DTDModels,Spitoni2021-TwoInfall} do not consider gas outflows and instead adjust the nucleosynthetic yields to reproduce the solar abundance. We adopt radially-dependent outflows as in \citetalias{Johnson2021-Migration} (see their Section 2.4 for details) for all our SFHs, including two-infall. We discuss the implications of this difference in Section \ref{sec:two-infall-discussion}.

@@

Similar trends can be seen when adjusting the timescale of the exponential DTD, as shown in the middle panel of Figure \ref{fig:onezone-threepanel}. Here, the knee is not a sharp feature associated with the onset of SNe Ia as in the power-law case, but rather a gentle curve in the abundance track around $t=1$ Gyr. Doubling of the timescale from 1.5 Gyr to 3 Gyr raises the [O/Fe] abundance ratio at $t=1$ Gyr by $\sim0.05$ dex and at $t=3$ Gyr by $\sim0.1$ dex, but the equilibrium abundance reached at $t=10$ Gyr remains unchanged. A longer exponential timescale also produces a broader [O/Fe] DF with more high-$\alpha$ stars, but the distribution is still unimodal.
The effect on the MDF is slightly more pronounced than the power-law case, with longer timescales skewing to lower [Fe/H] values.

@@

We run a total of multi-zone models with all combinations of our eight DTDs and four SFHs, for a total of 32. In the following subsections, we present the stellar abundance and age distributions from the multi-zone models and compare to APOGEE data from across the Galactic disk. \todo{Need to update plot descriptions and scores for runs with updated yields.}

@@

The abundance tracks of all four model SFHs initially follow a similar path of decreasing [O/Fe] with increasing [Fe/H]. The ISM abundance ratios of the inside-out model change monotonically over time. The stellar abundance distribution at both low- and high-[O/Fe] is composed of stars with a wide range of birth $R_{\rm gal}$. In particular, we observe a high-[O/Fe] ``ridge'' on the high-metallicity side of the ISM track, which is composed of stellar populations from the inner Galaxy.

@@

\begin{figure}
    \centering
    \includegraphics{figures/ofe_feh_twoinfall.pdf}
    \caption{The [O/Fe]--[Fe/H] plane for multiple Galactocentric regions from the model with the inside-out SFH and plateau ($W=1$ Gyr) DTD. The two columns of panels contain stars in different bins of $R_{\rm gal}$, and each row contains stars from a different bin of $|z|$. The contents of each panel is as described in Figure \ref{fig:ofe-feh-sfh}.}
    \label{fig:ofe-feh-twoinfall}
    \script{ofe_feh_twoinfall.py}
\end{figure}
